---
title: "Sarasota County Race Data Directory"
subtitle: "2020 Decennial Census race and ethnicity summary"
format:
  html:
    toc: true
    theme: journal
---

```{r}
#| echo: false
#| message: false
#| warning: false

library(tidyverse)

race_data <- read_csv("data/sarasota_race.csv")

race_data_long <- race_data |>
  pivot_longer(
    cols = c(total_pop, white, black, native_american, asian, pacific_islander, hispanic_latino),
    names_to = "category",
    values_to = "population"
  ) |>
  mutate(
    category = case_when(
      category == "total_pop" ~ "Total population",
      category == "white" ~ "White",
      category == "black" ~ "Black",
      category == "native_american" ~ "Native American",
      category == "asian" ~ "Asian",
      category == "pacific_islander" ~ "Pacific Islander",
      category == "hispanic_latino" ~ "Hispanic / Latino"
    )
  ) |>
  filter(category != "Total population")
```

## Overview

This website presents a formal summary of the Sarasota County race data directory, based on the 2020 Decennial Census dataset stored locally in the project. The page is intended to provide a concise and readable directory-style overview of the county's population composition.

### Summary

- Geographic area: `r race_data$name[1]`
- Total population: `r format(race_data$total_pop[1], big.mark = ",")`
- Source file: `data/sarasota_race.csv`

### Data table

```{r}
#| echo: false
#| message: false

race_data_long |>
  select(category, population) |>
  arrange(desc(population)) |>
  knitr::kable(
    col.names = c("Category", "Population"),
    align = c("l", "r")
  )
```

## Population by Race Group

```{r}
#| echo: false
#| warning: false

race_data_long |>
  ggplot(aes(x = category, y = population, fill = category)) +
  geom_col(width = 0.7) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Population by race group",
    x = "Race group",
    y = "Population"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold")
  )
```
